---
title: "DataClean"
author: 
- Miranda Corkum^1^
- Naaman Omar^1^
- Douglas Campbell^1^*
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: [RPackageCitations.bib, packages.bib]
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by the Canada Research Chairs.
NO was supported by 
DAC was supported by the Canada Research Chairs.


# Introduction

This script uses packages to combine marine microbial contamination data from DataStream (https://atlanticdatastream.ca) and weather data from Environment and Climate Change Canada (https://climate.weather.gc.ca/historical_data/search_historic_data_e.html) previously imported via API (Code/"DataStreamImport.Rmd", "WeatherCanImport.Rmd") at three sites on the Northumberland Strait: Shediac Bay, Kouchibouguac Bay and Parlee Beach Provincial Park. This combined bacteria/weather dataset (ProcessedData/"KouchShediacParleeData.Rds") is loaded into a Shiny app (Code/"Facet_app_WQ.R") to visualize water quality patterns. Based on those visualizations, statistical models and plots are then generated (see Code/"DataProcess.Rmd").

# Materials and Methods

## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
Set up libraries (or 'packages') - these contain additions to base R
Install all the packages in the tidyverse by running install. packages("tidyverse"). Run library(tidyverse) to load the core tidyverse and make it available in your current R session.
Note - lubridate is now part of Tidyverse version 2 or later.
```{r load libraries, include = FALSE}
library(tidyverse)

```

## Citations of software packages
Analysis was done using the 'tidyverse' [@R-tidyverse] package. 
Keep this chunk commented out or it will overwrite the final manuscript version of the .bib file.
```{r cite packages}
#knitr::write_bib(c('tidyverse'), 'packages.bib')

#To cite current version of R and R Studio:
#citation()
#RStudio.Version()

```

## Project Variables  
Assign project-specific values to variables
*Processed data contains any data that has been modified to include new columns etc. It can also contain the output of many files merged together.
*Tidied data contains a clean version of the raw data (Correct file names, dates in the right format etc).
*Do not use hard file path - not cross-platform compatible - use R functions to construct...
```{r set variables}
Project <- "northumberlandwaterquality"
DataIn <- file.path("..", "DataIn")
ProcessedData <- file.path("..", "ProcessedData")
Figures <- file.path("..", "Figures")
TidiedData <- file.path("..", "TidiedData")

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'

```

# Define bacterial thresholds for Health Canada Recreation (higher) and Shellfish Survey (lower)
```{r set thresholds}
threshold_1 <- 200
threshold_2 <- 43
```

## Kouchibouguac - Merge DataStream (bacteria) and ECCC (weather) datasets into a single file 
left_join will filter out from y any variable not in x (x dataframe is TidyData; y dataframe is KouchWeatherDay)
right_join does opposite of left_join
full_join will integrate all the data from y into x but will not filter out other data
Merge ECCC weather "date" column into DS "ActivityStartDate" column i.e. weather data only for dates with bacteria data

```{r read and join data kouch}
KouchData <- readRDS(file.path(TidiedData,"KouchData.Rds"))

KouchWeatherDay <- readRDS(file.path(TidiedData,"KouchWeatherDay.Rds"))

KouchDataAll <- KouchData %>% 
    full_join(x=., y=KouchWeatherDay, by = c("ActivityStartDate" = "date")) %>%
    filter(!is.na(DatasetName))

```

## Save new combined bacteria/weather data file for Kouchibouguac Bay
Q: Do I need to save these intermediary files? 
Q: Use locales here instead of hard path to place name?
```{r save kouch data}
saveRDS(KouchDataAll, file = file.path(ProcessedData, paste("KouchDataAll", ".Rds", sep = ""), fsep = .Platform$file.sep))
```

## Merge DataStream (bacteria) and ECCC (weather) datasets into a single file for Shediac
left_join will filter out from y any variable not in x (x dataframe is TidyData2; y dataframe is ShediacWeatherDay)
right_join - does opposite of left_join
full_join will integrate all the data from y into x but will not filter out other data

```{r read and join data shediac}
ShediacData <- readRDS(file.path(TidiedData,"ShediacData.Rds"))

ShediacWeatherDay <- readRDS(file.path(TidiedData,"ShediacWeatherDay.Rds"))

ShediacDataAll <- ShediacData %>% 
    full_join(x=., y=ShediacWeatherDay, by = c("ActivityStartDate" = "date")) %>%
    filter(!is.na(DatasetName))

```

## Save new combined bacteria/weather data file for Shediac Bay
```{r save Shediac data}
saveRDS(ShediacDataAll, file = file.path(ProcessedData, paste("ShediacDataAll", ".Rds", sep = ""), fsep = .Platform$file.sep))
```

## Merge DataStream (bacteria) and ECCC (weather) datasets into a single file for Parlee Beach Provincial Park (note: Moncton weather station data used for Shediac Bay is the same data for Parlee)
```{r read and join data parlee}
ParleeData <- readRDS(file.path(TidiedData,"ParleeData.Rds"))

ParleeDataAll <- ParleeData %>% 
    full_join(x=., y=ShediacWeatherDay, by = c("ActivityStartDate" = "date")) %>%
    filter(!is.na(DatasetName))

```

## Save new combined bacteria/weather data file for Parlee Beach Provincial Park
```{r save Parlee data}
saveRDS(ParleeDataAll, file = file.path(ProcessedData, paste("ParleeDataAll", ".Rds", sep = ""), fsep = .Platform$file.sep))
```

## 1) Combine KouchData, ShediacData and Parlee Data to obtain a single dataframe to use in Shiny app data exploration
## 2) Remove NA columns - drop columns only AFTER merging data, otherwise the dataframes will not bind
Drop from 80 to 65 columns
https://community.rstudio.com/t/drop-all-na-columns-from-a-dataframe/5844/6
## 3) Add column for Julian Day = 66 columns
## 4) Add column for Health Canada bacteria limit = 67 columns

Add lines to deal with Result Values below limit of detection:
1) If ResultValue is zero, report Bacterial_Count as "1";
2) If ResultValue is NA AND ResultDetectionQuantitationMeasure is 2, report Bacterial_Count as "1";
3) If ResultValue is NA AND ResultDetectionQuantitationMeasure is 10, report Bacterial_Count as "5";
4) If ResultValue is NA AND ResultDetectionQuantitationMeasure is 400, report Bacterial_Count as "400".

```{r bind dataframes}
KouchShediacParleeData <- rbind(KouchDataAll, ShediacDataAll, ParleeDataAll)
KouchShediacParleeData <- KouchShediacParleeData %>%
  discard(~all(is.na(.x))) %>%
  map_df(~.x) %>%
  mutate(JulianDay = yday(ActivityStartDate)) %>%
  mutate(ResultValueColour = case_when(ResultValue >= threshold_1 ~ "red",
                               ResultValue >= threshold_2 ~ "blue",
                               ResultValue < threshold_2 ~ "black")) %>%
  mutate(MonitoringLocationNameStrip = case_when(MonitoringLocationName == "Kouchibouguac Bay" ~ "Kouchibouguac Bay", 
                                                 MonitoringLocationName == "Shediac Bay" ~ "Shediac Bay",
                                                 MonitoringLocationName == "Parlee Beach Provincial Park" ~ "Parlee Beach")) %>%
  rename(Measurement_Date = ActivityStartDate) %>%
  mutate(Bacterial_Count = case_when(is.na(ResultValue) & ResultDetectionQuantitationLimitMeasure == 2 ~ 1,
                                     is.na(ResultValue) & ResultDetectionQuantitationLimitMeasure == 10 ~ 5,
                                     is.na(ResultValue) & ResultDetectionQuantitationLimitMeasure == 400 ~ 400,
                                     ResultValue == 0 ~ 1,
                                     TRUE~ResultValue)) 
  
#names(KouchShediacParleeData)

```

# In order for facet plots to be stacked in correct order in DataProcess.Rmd we need to first reorder station names in new column "MonitoringLocationNameStrip".
```{r data reorder}
KouchShediacParleeData$MonitoringLocationNameStrip <- factor(KouchShediacParleeData$MonitoringLocationNameStrip, levels=c('Kouchibouguac Bay', 'Shediac Bay', 'Parlee Beach'))
```


```{r save all data}
saveRDS(KouchShediacParleeData, file = file.path(ProcessedData, paste("KouchShediacParleeData", ".Rds", sep = ""), fsep = .Platform$file.sep))
#write.csv(KouchShediacParleeData, file = file.path(ProcessedData, paste("KouchShediacParleeData", ".csv", sep = ""), fsep = .Platform$file.sep), row.names=FALSE)

```
