---
title: "DataProcess"
author: 
- Miranda Corkum^1^
- Naaman Omar^1^
- Douglas Campbell^1^*
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: [RPackageCitations.bib, packages.bib]
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by the Canada Research Chairs.
NO was supported by 
DAC was supported by the Canada Research Chairs.


# Introduction

This script takes the combined bacteria/weather data file generated in "DataClean.Rmd" (ProcessedData/"KouchShediacParleeData.Rds") and applies a binomial model to test for interactions between variables. Various tables and figures are generated for use in the final manuscript "NorthumberlandManuscript.Rmd" (Figures/"StatsAll_table1.rds", "StatsAll_table2.rds", "KouchShediacParleeBinom_fig.png", "StatsAll_table3.rds", "precip_prior48_fig.png", "measurement_date_fig.png", "mean_temp_prior48_fig.png").

# Materials and Methods

## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
Set up libraries (or 'packages') - these contain additions to base R
Install all the packages in the tidyverse by running install. packages("tidyverse"). Run library(tidyverse) to load the core tidyverse and make it available in your current R session.
Note - lubridate is now part of Tidyverse version 2 or later.
```{r load libraries, include = FALSE}
library(tidyverse)

#required to run binomial model - converts statistical objects into Tidy Tibbles
library(broom)

#required for density plots
library(ggpubr)

#required for tables
library(kableExtra)

#required for plot colour scaling
library(viridis)  

```

```{r logit2prob, include = FALSE}
source("https://sebastiansauer.github.io/Rcode/logit2prob.R")
```

## Citations of software packages
Analysis was done using the 'tidyverse' [@R-tidyverse], 'broom' [@R-broom], 'ggpubr' [@R-ggpubr], 'kableExtra' [@R-kableExtra] and 'viridis' [@R-viridis] packages. 
Keep this chunk commented out or it will overwrite the final manuscript version of the .bib file.
```{r cite packages}
#knitr::write_bib(c('tidyverse', 'broom', 'ggpubr', 'kableExtra', 'viridis'), 'packages.bib')

#To cite current version of R and R Studio:
#citation()
#RStudio.Version()

```

## Project Variables  
Assign project-specific values to variables

```{r set variables}
Project <- "northumberlandwaterquality"
DataIn <- file.path("..", "DataIn")
ProcessedData <- file.path("..", "ProcessedData")
Figures <- file.path("..", "Figures")
TidiedData <- file.path("..", "TidiedData")

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'

```

# Define bacterial thresholds for Health Canada Recreation (higher) and Shellfish Survey (lower)
```{r set thresholds}
threshold_1 <- 200
threshold_2 <- 43
```


```{r read .Rds}
KouchShediacParleeData <- readRDS(file.path(ProcessedData,"KouchShediacParleeData.Rds"))

```
9May2023
Notes from Naaman - binomial model
Create a new column with a logical vector = above/below threshold
Result_Value_logical
  mutate piped into as.logical (want only 0s or 1s)
  can also look for covariance i.e. an analysis of covariance with the slopes. This would tell us whether the slopes are significantly different.
  logit - an inverse log
  tidy - function within broom
  map - works with nested things
  glm = generalized linear model
  lose statistical power running model on bins vs raw data
Binomial distribution is a statistical probability distribution that states the likelihood that a value will take one of two independent values under a given set of parameters or assumptions.
Caption: Bacterial_Count vs precipitation with mean temperature as covariant (i.e. effects of temp have been factored in).

Note: "+" means non-interacting terms; "*" means interacting terms

## Create binomial model 
24Aug2023
Changed the "+" to "*" to test for interactions between year and precip. We see an interaction for Parlee only. i.e. Different years had different amount of rain in Parlee. For Parlee, we can't know if precip or different years are causing bacterial counts.
- Break model up into individual factors: precip still significant at Parlee and Shediac. Temp is not sig. at any location. Therefore, there is no effect of temperature on bacterial counts. Bacteria is not simply present and growing in the water at these beaches during warmer months, rather it is getting washed in.
- Interaction factor loses a degree of freedom (= less statistical power).
- Use final precip_year model; update table to include this version.

Note to Felix: Did we test for colinearity? No, but this is not required for a binomial model. For a binomial model we need to test for interactions (*) instead, which we have done.
12Sept2023 TEST: "precip_prior48" = precip_lag24 + precip_lag48
6Oct2023 TEST: 
use ECCC threshold_1 i.e. 200 MPN
Q for Naaman - why aren't lines straight?

23Jan2024
Update model to use prior48 values for temp and precip
Now we see a significant effect of temp for Parlee but not Kouch or Shediac
SO maybe switch to using precip year temp binomial model (instead of precip year only binomial model)


## MODEL 1 - Bacterial threshold_1 = 200 MPN/100 mL
"Bacterial_Count" vs "precip_sum72" OR precip_prior48 covarying (non-interacting) with "year" AND "mean_temp_AVG72" 
```{r binomial precip year temp}
KouchShediacParleeBinom_1 <- KouchShediacParleeData %>%
  mutate(Bacterial_Count_logical = case_when(Bacterial_Count >= threshold_1 ~1, 
                                         Bacterial_Count < threshold_1 ~ 0),
         Bacterial_Count_logical = Bacterial_Count_logical %>% as.logical() %>% as.numeric()) %>%
  nest(NestedData = -c(MonitoringLocationNameStrip)) %>%
  mutate(precip_year_temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year) + .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_temp_binomial_model_tidy = map(precip_year_temp_binomial_model, tidy),
         precip_year_temp_binomial_model_augment = map(precip_year_temp_binomial_model, augment)) %>%
   mutate(precip_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_binomial_model_tidy = map(precip_binomial_model, tidy),
         precip_binomial_model_augment = map(precip_binomial_model, augment)) %>%
     mutate(temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         temp_binomial_model_tidy = map(temp_binomial_model, tidy),
         temp_binomial_model_augment = map(temp_binomial_model, augment)) %>%
       mutate(year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         year_binomial_model_tidy = map(year_binomial_model, tidy),
         year_binomial_model_augment = map(year_binomial_model, augment)) %>%
         mutate(precip_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_binomial_model_tidy = map(precip_year_binomial_model, tidy),
         precip_year_binomial_model_augment = map(precip_year_binomial_model, augment)) %>%
         mutate(precip48_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip48_year_binomial_model_tidy = map(precip48_year_binomial_model, tidy),
         precip48_year_binomial_model_augment = map(precip48_year_binomial_model, augment)) 

```

## Create tables

## Table Model 1
precip 48 h prior and year with Bacterial threshold_1 = 200 MPN/100 mL
```{r bind table precip 48 prior year 200 MPN}

StatsAll = data.frame()

for (index in 1:3) {
  StatsAll <- rbind(StatsAll, KouchShediacParleeBinom_1$precip_year_temp_binomial_model_tidy[[index]] %>%
    select(-c(std.error,statistic)) %>%
    filter(term != "(Intercept)") %>%
    pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
    mutate(Location = KouchShediacParleeBinom_1$MonitoringLocationNameStrip[index], .before = `estimate_.$precip_prior48`) %>%
    rename(Precip_prior48 = `estimate_.$precip_prior48`,
           Year = `estimate_as.numeric(.$year)`,
           Temp_prior48 = `estimate_.$mean_temp_AVG48`,
           Precip_prior48_pvalue = `p.value_.$precip_prior48`,
           Year_pvalue = `p.value_as.numeric(.$year)`,
           Temp_pvalue = `p.value_.$mean_temp_AVG48`))
}

StatsAll_table1 <- StatsAll

kable_styling(kable(StatsAll_table1, digits = 3), position = "left")

```

## Create plots

In the face of recent (2017) public water quality concerns, management changes and increased water monitoring at Parlee Beach have led to fewer days with bacterial counts above threshold. 

Explaining the line on model plot: the binomial probability of a given count on a given day being above threshold given the covarying terms of cumulative precipitation and mean temperature over 72 hours.
DC - set colour scale for year to contrast and show that later years are decreasing
Naaman - use hex codes or colour names
13Sept - Define "probability above threshold" in figure caption.

+  scale_color_viridis(discrete = TRUE, direction = -1, option = "D") +

WHy aren't lines straight?
logit = ln(x/1-x)
Like if you invert a sigmoid curve!

## Plot Model 1
```{r plot model precip_prior48 year 200 MPN}

KouchShediacParleeBinom_1_fig <- KouchShediacParleeBinom_1 %>%
  unnest(precip48_year_binomial_model_tidy) %>%
  unnest(precip48_year_binomial_model_augment) %>% 
  mutate(Year = as.character(`as.numeric(.$year)`)) %>% 
  ggplot() +
  geom_point(aes(x = `.$precip_prior48`, y = `.$Bacterial_Count_logical`, colour = Year)) +
  geom_line(aes(x = `.$precip_prior48`, y = logit2prob(.fitted), colour = Year)) +
  geom_text(aes(x = 25, y = 0.6, label = paste("pvalue = ", signif(`p.value`, 2))), data = .%>% filter(term == ".$precip_prior48")) +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme_bw() +
  theme(strip.text.y = element_text(size = 10),
        axis.title = element_text(size = 14)) +
  scale_color_viridis(discrete = TRUE, direction = -1, option = "C") +
  #scale_colour_manual(values = c("#FFF59D", "#FFF176", "#FDD835", "#FFC107", "#F39C12", "#F57C00", "#E64A19", "#D32F2F", "#FF0000", "#E5E7E9", "#BDC3C7", "#4D5656", "#17202A")) +
  #scale_colour_manual(values = c("#000000", "#607D8B", "#33CC00", "#33FF00", "#99FF00", "#99CC00", "#993300", "#FF0000", "#FF6600", "#FF9900", "#FFCC00", "#FFFF00")) +  
    xlab("Cumulative Precipitation (mm) over 48 h") + ylab("Probability Above Threshold")

KouchShediacParleeBinom_1_fig
```


## MODEL 2
Bacterial threshold_2 = 43 MPN/100 mL 
```{r binomial precip year temp}
KouchShediacParleeBinom_2 <- KouchShediacParleeData %>%
  mutate(Bacterial_Count_logical = case_when(Bacterial_Count >= threshold_2 ~1, 
                                         Bacterial_Count < threshold_2 ~ 0),
         Bacterial_Count_logical = Bacterial_Count_logical %>% as.logical() %>% as.numeric()) %>%
  nest(NestedData = -c(MonitoringLocationNameStrip)) %>%
  mutate(precip_year_temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year) + .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_temp_binomial_model_tidy = map(precip_year_temp_binomial_model, tidy),
         precip_year_temp_binomial_model_augment = map(precip_year_temp_binomial_model, augment)) %>%
   mutate(precip_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_binomial_model_tidy = map(precip_binomial_model, tidy),
         precip_binomial_model_augment = map(precip_binomial_model, augment)) %>%
     mutate(temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         temp_binomial_model_tidy = map(temp_binomial_model, tidy),
         temp_binomial_model_augment = map(temp_binomial_model, augment)) %>%
       mutate(year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         year_binomial_model_tidy = map(year_binomial_model, tidy),
         year_binomial_model_augment = map(year_binomial_model, augment)) %>%
         mutate(precip_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_binomial_model_tidy = map(precip_year_binomial_model, tidy),
         precip_year_binomial_model_augment = map(precip_year_binomial_model, augment)) %>%
         mutate(precip48_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip48_year_binomial_model_tidy = map(precip48_year_binomial_model, tidy),
         precip48_year_binomial_model_augment = map(precip48_year_binomial_model, augment)) 


```

## Table Model 2
supplemental table for precip 48 h prior with Bacterial threshold_2 = 43 MPN/100 mL

```{r bind table precip 48 prior year 43 MPN}

StatsAll = data.frame()

for (index in 1:3) {
  StatsAll <- rbind(StatsAll, KouchShediacParleeBinom_2$precip_year_temp_binomial_model_tidy[[index]] %>%
    select(-c(std.error,statistic)) %>%
    filter(term != "(Intercept)") %>%
    pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
    mutate(Location = KouchShediacParleeBinom_2$MonitoringLocationNameStrip[index], .before = `estimate_.$precip_prior48`) %>%
    rename(Precip_prior48 = `estimate_.$precip_prior48`,
          Year = `estimate_as.numeric(.$year)`,
          Temp_prior48 = `estimate_.$mean_temp_AVG48`,
          Precip_prior48_pvalue = `p.value_.$precip_prior48`,
          Year_pvalue = `p.value_as.numeric(.$year)`,
          Temp_pvalue = `p.value_.$mean_temp_AVG48`))
}

StatsAll_table2 <- StatsAll

kable_styling(kable(StatsAll_table2, digits = 3), position = "left")

```

## Plot Model 2
```{r plot model precip_prior48 year 43 MPN}
KouchShediacParleeBinom_2_fig <- KouchShediacParleeBinom_2 %>%
  unnest(precip48_year_binomial_model_tidy) %>%
  unnest(precip48_year_binomial_model_augment) %>% 
  mutate(Year = as.character(`as.numeric(.$year)`)) %>% 
  ggplot() +
  geom_point(aes(x = `.$precip_prior48`, y = `.$Bacterial_Count_logical`, colour = Year)) +
  geom_line(aes(x = `.$precip_prior48`, y = logit2prob(.fitted), colour = Year)) +
  geom_text(aes(x = 25, y = 0.6, label = paste("pvalue = ", signif(`p.value`, 2))), data = .%>% filter(term == ".$precip_prior48")) +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme_bw() +
  theme(strip.text.y = element_text(size = 10),
        axis.title = element_text(size = 14)) +
  scale_color_viridis(discrete = TRUE, direction = -1, option = "C") +
  #scale_colour_manual(values = c("#FFF59D", "#FFF176", "#FDD835", "#FFC107", "#F39C12", "#F57C00", "#E64A19", "#D32F2F", "#FF0000", "#E5E7E9", "#BDC3C7", "#4D5656", "#17202A")) +
  #scale_colour_manual(values = c("#000000", "#607D8B", "#33CC00", "#33FF00", "#99FF00", "#99CC00", "#993300", "#FF0000", "#FF6600", "#FF9900", "#FFCC00", "#FFFF00")) +  
    xlab("Cumulative Precipitation (mm) over 48 h") + ylab("Probability Above Threshold")

KouchShediacParleeBinom_2_fig
```

## MODEL 3
Test model rain days with < 40 mm (i.e. precip_prior48 values < 40) with Bacterial threshold_1 = 200 MPN/100 mL
Filter data - line count drops from 7233 to 6904 observations.
```{r filter precip_prior48}
KouchShediacParleeData_40 <- filter(KouchShediacParleeData, precip_prior48 < 40) 
```


```{r binomial precip year temp}
KouchShediacParleeBinom_3 <- KouchShediacParleeData_40 %>%
  mutate(Bacterial_Count_logical = case_when(Bacterial_Count >= threshold_1 ~1, 
                                         Bacterial_Count < threshold_1 ~ 0),
         Bacterial_Count_logical = Bacterial_Count_logical %>% as.logical() %>% as.numeric()) %>%
  nest(NestedData = -c(MonitoringLocationNameStrip)) %>%
 mutate(precip_year_temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year) + .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_temp_binomial_model_tidy = map(precip_year_temp_binomial_model, tidy),
         precip_year_temp_binomial_model_augment = map(precip_year_temp_binomial_model, augment)) %>%
   mutate(precip_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_binomial_model_tidy = map(precip_binomial_model, tidy),
         precip_binomial_model_augment = map(precip_binomial_model, augment)) %>%
     mutate(temp_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$mean_temp_AVG48, data = ., family = binomial(link = "logit")), otherwise = NULL)),
         temp_binomial_model_tidy = map(temp_binomial_model, tidy),
         temp_binomial_model_augment = map(temp_binomial_model, augment)) %>%
       mutate(year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         year_binomial_model_tidy = map(year_binomial_model, tidy),
         year_binomial_model_augment = map(year_binomial_model, augment)) %>%
         mutate(precip_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip_year_binomial_model_tidy = map(precip_year_binomial_model, tidy),
         precip_year_binomial_model_augment = map(precip_year_binomial_model, augment)) %>%
         mutate(precip48_year_binomial_model = map(NestedData, possibly(~glm(.$Bacterial_Count_logical ~ .$precip_prior48 + as.numeric(.$year), data = ., family = binomial(link = "logit")), otherwise = NULL)),
         precip48_year_binomial_model_tidy = map(precip48_year_binomial_model, tidy),
         precip48_year_binomial_model_augment = map(precip48_year_binomial_model, augment)) 


```

## Table Model 3
precip 48 h prior < 40 mm and year with Bacterial threshold_1 = 200 MPN/100 mL (Model 3)
```{r bind table precip 48 prior year 200 MPN less 40}

StatsAll = data.frame()

for (index in 1:3) {
  StatsAll <- rbind(StatsAll, KouchShediacParleeBinom_3$precip_year_temp_binomial_model_tidy[[index]] %>%
    select(-c(std.error,statistic)) %>%
    filter(term != "(Intercept)") %>%
    pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
    mutate(Location = KouchShediacParleeBinom_3$MonitoringLocationNameStrip[index], .before = `estimate_.$precip_prior48`) %>%
    rename(Precip_prior48 = `estimate_.$precip_prior48`,
          Year = `estimate_as.numeric(.$year)`,
          Temp_prior48 = `estimate_.$mean_temp_AVG48`,
          Precip_prior48_pvalue = `p.value_.$precip_prior48`,
          Year_pvalue = `p.value_as.numeric(.$year)`,
          Temp_pvalue = `p.value_.$mean_temp_AVG48`))
}

StatsAll_table3 <- StatsAll

kable_styling(kable(StatsAll_table3, digits = 3), position = "left")

```

## Plot Model 3
```{r plot model precip_prior48 year 200 MPN less 40}
KouchShediacParleeBinom_3_fig <- KouchShediacParleeBinom_3 %>%
  unnest(precip48_year_binomial_model_tidy) %>%
  unnest(precip48_year_binomial_model_augment) %>% 
  mutate(Year = as.character(`as.numeric(.$year)`)) %>% 
  ggplot() +
  geom_point(aes(x = `.$precip_prior48`, y = `.$Bacterial_Count_logical`, colour = Year)) +
  geom_line(aes(x = `.$precip_prior48`, y = logit2prob(.fitted), colour = Year)) +
  geom_text(aes(x = 25, y = 0.6, label = paste("pvalue = ", signif(`p.value`, 2))), data = .%>% filter(term == ".$precip_prior48")) +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme_bw() +
  theme(strip.text.y = element_text(size = 10),
        axis.title = element_text(size = 14)) +
  scale_color_viridis(discrete = TRUE, direction = -1, option = "C") +
  #scale_colour_manual(values = c("#FFF59D", "#FFF176", "#FDD835", "#FFC107", "#F39C12", "#F57C00", "#E64A19", "#D32F2F", "#FF0000", "#E5E7E9", "#BDC3C7", "#4D5656", "#17202A")) +
  #scale_colour_manual(values = c("#000000", "#607D8B", "#33CC00", "#33FF00", "#99FF00", "#99CC00", "#993300", "#FF0000", "#FF6600", "#FF9900", "#FFCC00", "#FFFF00")) +  
    xlab("Cumulative Precipitation (mm) over 48 h") + ylab("Probability Above Threshold")

KouchShediacParleeBinom_3_fig
```


# png(file = file.path(Figures, paste("StatsAll_table2", ".png", sep = ""), fsep = .Platform$file.sep), height=200, width=2500)
# p <- tableGrob(StatsAll_table2, rows=NULL, theme = ttheme_default(base_size = 40))
# grid.arrange(p)
# dev.off()


## General data plots
precip is Poisson distributed because majority of days we have no rain (mean near 0), negative values not possible.

```{r ggplot precip_prior48}
precip_prior48_fig <- KouchShediacParleeData %>%   
  ggplot() + 
  geom_point(aes(x = precip_prior48, y = Bacterial_Count, colour = ResultValueColour)) +
  scale_y_continuous(trans='log10') +
  scale_x_continuous(breaks=seq(0,125,10)) +
  #scale_y_continuous(aes(breaks = seq(0, max(ResultValue), by = 0.1))) +
  annotation_logticks(sides = "l", colour = "gray60") +
  geom_hline(yintercept = 200, linetype='dotted', col = 'red') +
  geom_hline(yintercept = 43, linetype='dotted', col = 'blue') +
  scale_colour_identity() +
  theme_bw() +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme(strip.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) +
  xlab("Cumulative Precipitation (mm) over 48 h") + ylab("Bacterial Count ("~log[10] ~ "scale)")

precip_prior48_fig
```


```{r ggplot date}
measurement_date_fig <- KouchShediacParleeData %>%   
  ggplot() + 
  geom_point(aes(x = Measurement_Date, y = Bacterial_Count, colour = ResultValueColour)) +
  scale_x_datetime(date_breaks = "1 year", date_labels="%Y") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l", colour = "gray60") +
  geom_hline(yintercept = 200, linetype='dotted', col = 'red') +
  geom_hline(yintercept = 43, linetype='dotted', col = 'blue') +
  scale_colour_identity() +
  theme_bw() +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme(strip.text.y = element_text(size = 10)) +
  theme(axis.title = element_text(size = 14)) +
  xlab("Sample Collection Date") + ylab("Bacterial Count ("~log[10] ~ "scale)")

measurement_date_fig
```
# Shapiro-Wilk test
If the p-value of the Shapiro-Wilk Test is greater than 0.05, the data is normal. If it is below 0.05, the data significantly deviate from a normal distribution.

```{r Shapiro-Wilk Test Kouch}
KouchShapiro <- KouchShediacParleeData %>%
  filter(MonitoringLocationName == "Kouchibouguac Bay") 

shapiro.test(KouchShapiro$mean_temp_prior48) %>% 
  tidy() %>% 
  pluck("p.value")
```
```{r Shapiro-Wilk Test Shediac}
ShediacShapiro <- KouchShediacParleeData %>%
  filter(MonitoringLocationName == "Shediac Bay") 

shapiro.test(ShediacShapiro$mean_temp_prior48) %>% 
  tidy() %>% 
  pluck("p.value")
```

```{r Shapiro-Wilk Test Parlee}
ParleeShapiro <- KouchShediacParleeData %>%
  filter(MonitoringLocationName == "Parlee Beach Provincial Park") 

shapiro.test(ParleeShapiro$mean_temp_prior48) %>% 
  tidy() %>% 
  pluck("p.value")
```
```{r density}
#KouchShediacParleeData %>% 
  #ggdensity(.$mean_temp_prior48,
          #xlab = "Air Temperature (°C) over 48 h") +
  #facet_grid(rows = vars(MonitoringLocationName))

KouchShediacParleeData %>% 
  ggplot() + 
  geom_density(aes(x = mean_temp_prior48, fill = MonitoringLocationName), alpha = 0.3)

```
Data are left skewed - tail on the left.
Data is not normally distributed.


```{r ggplot mean temp prior48}
mean_temp_prior48_fig <- KouchShediacParleeData %>%
  mutate(CharacteristicName = case_when(CharacteristicName == "Escherichia coli" ~ "Fecal Coliform", TRUE ~ CharacteristicName)) %>% 
  ggplot() +
  geom_point(aes(x = mean_temp_AVG72, y = Bacterial_Count, colour = ResultValueColour)) +
  scale_y_continuous(trans='log10') +
  scale_x_continuous(breaks=seq(0,27,5)) +
  annotation_logticks(sides = "l", colour = "gray60") +
  geom_hline(yintercept = 200, linetype='dotted', col = 'red') +
  geom_hline(yintercept = 43, linetype='dotted', col = 'blue') +
  scale_colour_identity() +
  theme_bw() +
  facet_grid(rows = vars(MonitoringLocationNameStrip)) +
  theme(strip.text.y = element_text(size = 10)) +
  scale_shape_manual(values = c(19, 8)) +
  theme(axis.title = element_text(size = 14)) +
  labs(x= "Air Temperature (°C) over 48 h",
       y = "Bacterial Count ("~log[10] ~ "scale)")


mean_temp_prior48_fig
```
Group: even if there are multiple lines within a year (because of different temps) return them separately.
Temperature has an impact on probability of counts being above a threshold for Parlee (43 and 200 MPN) and for Shediac Bay (43 MPN only).


```{r}
KouchShediacParleeBinom_1 <- KouchShediacParleeBinom_1 %>% 
  mutate(Threshold = threshold_1) 

KouchShediacParleeBinom_2 <- KouchShediacParleeBinom_2 %>% 
  mutate(Threshold = threshold_2) 

KouchShediacParleeBinom_Combined <- rbind(KouchShediacParleeBinom_1, KouchShediacParleeBinom_2)

KouchShediacParleeBinom_fig <- KouchShediacParleeBinom_Combined %>% 
  unnest(precip_year_temp_binomial_model_tidy) %>%
  unnest(precip_year_temp_binomial_model_augment) %>% 
  mutate(Year = as.character(`as.numeric(.$year)`)) %>% 
  mutate(striplabel = "Threshold",
         countsmL = case_when(Threshold == 43 ~ "43 counts per mL",
                              Threshold == 200 ~ "200 counts per mL"), 
         countsmL = factor(countsmL, levels = c("43 counts per mL", "200 counts per mL"))) %>% 
  ggplot() +
  geom_point(aes(x = `.$precip_prior48`, y = `.$Bacterial_Count_logical`, colour = Year)) +
  geom_line(aes(x = `.$precip_prior48`, y = logit2prob(.fitted), colour = Year, group = `.$mean_temp_AVG48`)) +
  geom_text(aes(x = 50, y = 0.75, label = paste("p value = ", signif(`p.value`, 2))), data = .%>% filter(term == ".$precip_prior48")) +
  ggh4x::facet_nested(rows = vars(MonitoringLocationNameStrip), cols = vars(striplabel, countsmL), labeller = labeller(MonitoringLocationNameStrip = label_value, countsmL = label_value)) +
  theme_bw() +
  theme(strip.text.y = element_text(size = 9),
        axis.title = element_text(size = 14)) +
  scale_color_viridis(discrete = TRUE, direction = -1, option = "C") +
  xlab("Cumulative Precipitation (mm) over 48 h") + ylab("Probability of Counts Above Threshold") 

KouchShediacParleeBinom_fig

```



```{r save ggplots and tables}

saveRDS(object = StatsAll_table1, file = file.path("..", "Figures", "StatsAll_table1.rds"))

saveRDS(object = StatsAll_table2, file = file.path("..", "Figures", "StatsAll_table2.rds"))

saveRDS(object = StatsAll_table3, file = file.path("..", "Figures", "StatsAll_table3.rds"))

ggsave(paste("KouchShediacParleeBinom_fig",".png",sep = ""), plot = KouchShediacParleeBinom_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("Fig5",".tiff",sep = ""), plot = KouchShediacParleeBinom_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("precip_prior48_fig",".png",sep = ""), plot = precip_prior48_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("Fig4",".tiff",sep = ""), plot = precip_prior48_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("measurement_date_fig",".png",sep = ""), plot = measurement_date_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("Fig2",".tiff",sep = ""), plot = measurement_date_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("mean_temp_prior48_fig",".png",sep = ""), plot = mean_temp_prior48_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

ggsave(paste("Fig3",".tiff",sep = ""), plot = mean_temp_prior48_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

#ggsave(paste("KouchShediacParleeBinom_1_fig",".png",sep = ""), plot = KouchShediacParleeBinom_1_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

#ggsave(paste("KouchShediacParleeBinom_2_fig",".png",sep = ""), plot = KouchShediacParleeBinom_2_fig, device = NULL, path = Figures,scale = 1, height=5, #width= 7, units = c("in"),dpi = 300, limitsize = TRUE)

#ggsave(paste("KouchShediacParleeBinom_3_fig",".png",sep = ""), plot = KouchShediacParleeBinom_3_fig, device = NULL, path = Figures,scale = 1, height=5, width= 7, units = c("in"),dpi = 300, limitsize = TRUE)
```
