---
title: "ShellfishSurvey_DSConvert"
author: 
- Miranda Corkum^1^
- Naaman Omar^1^
- Douglas Campbell^1^*
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: [RPackageCitations.bib, packages.bib]
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by the Canada Research Chairs.
NO was supported by 
DAC was supported by the Canada Research Chairs.


# Introduction

This script uses packages to import and format Environment and Climate Change Canada Shellfish Survey data from "Shellfish Water Classification Program – Marine Water Quality Data in New Brunswick" for years 2016-2022. Data is downloaded manually as .csv from the Open Government Data Portal as two separate files a) bacteria data: DataIn/"Shellfish_Samples_Mollusques_Echantillons_NB.csv" and b) monitoring location data: DataIn/"Shellfish_Sites_Mollusques_NB.csv". These file are then imported into R for tidying and conversion to DataStream upload format (TidiedData/"ECCCDataKouchShediac.Rds"). This new data is added to the existing DataStream dataset and further manipulated in Code/"DataStreamImport.Rmd". Note: the Kouchibouguac Bay and Shediac Bay datasets targeted for update in this script were not uploaded to DataStream because the DataStream team are currently working to implement an automated direct data download process from the ECCC Shellfish Survey database.

# Materials and Methods

## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
Set up libraries (or 'packages') - these contain additions to base R
Install all the packages in the tidyverse by running install. packages("tidyverse"). Run library(tidyverse) to load the core tidyverse and make it available in your current R session.
```{r load libraries, include = FALSE}
library(tidyverse)

```

## Citations of software packages
Analysis was done using the 'tidyverse' [@R-tidyverse] package. 
Keep this chunk commented out or it will overwrite the final manuscript version of the .bib file.
```{r cite packages}
#knitr::write_bib(c('tidyverse'), 'packages.bib')
```

## Project Variables  
Assign project-specific values to variables

```{r set variables}
Project <- "northumberlandwaterquality"
DataIn <- file.path("..", "DataIn")
ProcessedData <- file.path("..", "ProcessedData")
Figures <- file.path("..", "Figures")
TidiedData <- file.path("..", "TidiedData")

FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'

```

## Import two data objects previously downloaded from ECCC Open Data Portal (NB Shellfish Survey a) data and b) station locations) and stored as .csv
```{r read csv}
ECCCData <- read_csv(file = file.path(DataIn, "Shellfish_Samples_Mollusques_Echantillons_NB.csv", fsep =.Platform$file.sep))
ECCCLocations <- read_csv(file = file.path(DataIn, "Shellfish_Sites_Mollusques_NB.csv", fsep =.Platform$file.sep))
```

## Save the imported data as .Rds for further data exploration (allows for rapid R import)
```{r save rds}
saveRDS(ECCCData, file = file.path(DataIn, paste("ECCCData", ".Rds", sep = ""), fsep = .Platform$file.sep))
saveRDS(ECCCLocations, file = file.path(DataIn, paste("ECCCLocations", ".Rds", sep = ""), fsep = .Platform$file.sep))

```

*Test with Kouch data entries only first
## Subset data - filter for Kouchibouguac Bay (Sector ID "NB-05-010-003") data from year 2016 onwards; 21 sites in total (some not active)
322 lines

```{r subset data kouch}
ECCCData2 <- filter(ECCCData, Sector_Secteur == "NB-05-010-003" & Date >= "2016-01-01") 
```

## Combine and re-format columns "Sector_Secteur" (i.e. NB-05-010-003) and "Site" (i.e. NB_0019) for both the data and site location files.
```{r}
ECCCData2$Site <- gsub('[NB_]', '', ECCCData2$Site)

ECCCData2$MonitoringLocationID <- paste(ECCCData2$Sector_Secteur, ECCCData2$Site)

ECCCLocations$Site <- gsub('[NB_]', '', ECCCLocations$Site)

ECCCLocations$MonitoringLocationID <- paste(ECCCLocations$Sector_Secteur, ECCCLocations$Site)

```

## Join two dataframes by matching "Sector_Secteur" and "Site" columns.
```{r join dataframes}
ECCCDataKouch <- ECCCData2 %>% 
    full_join(x=., y=ECCCLocations, by = c("MonitoringLocationID" = "MonitoringLocationID")) %>%
    filter(!is.na(Date))
```

## Select relevant columns - start with bacteria only, include temp and salinity data later on.
```{r column select}
ECCCDataKouch <- ECCCDataKouch %>%
  select(MonitoringLocationID, Latitude, Longitude, Date, FC_MPN_CF_NPP_100_mL) 

#select(MonitoringLocationID, Latitude, Longitude, Date, FC_MPN_CF_NPP_100_mL, Temp_C, Sal_PPT_PPM) 
```

## Rename current columns to DataStream format
```{r column rename}
ECCCDataKouch <- ECCCDataKouch %>%
  rename("ResultValue" = "FC_MPN_CF_NPP_100_mL", "MonitoringLocationLatitude" = "Latitude", "MonitoringLocationLongitude" = "Longitude", "ActivityStartDate" = "Date")

#Rename temp and salinity data later on:
#rename("Fecal Coliform" = "FC_MPN_CF_NPP_100_mL", "Temperature, water" = "Temp_C", "Salinity" = "Sal_PPT_PPM", "ActivityStartDate" = "Date")
```

## Use pivot longer only if including salinity and temperature data; exclude at first to simplify
```{r pivot longer}
# ECCCDataAll %>%
#   pivot_longer(
#     #cols = c("Fecal Coliform", "Temperature, water", "Salinity"),
#     names_to = "CharacteristicName",
#     values_to = "ResultValue"
#   )

```

## Date format
Tell R to recognize the date column as "date"
```{r date format}
ECCCDataKouch <- tidyr::separate(ECCCDataKouch, ActivityStartDate, c("ActivityStartDate", "ActivityStartTime"), sep = " ") %>%
  mutate(ActivityStartDate = ymd(ActivityStartDate))
```

## Add all DataStream columns
Will need to repeat this chunk for each new location (i.e. "Shediac Bay") or parameter (i.e. temp and salinity) to be added. If adding many locations, should write a function instead. For now, adding only 2 locations - this way okay.
```{r add columns}
#define new columns to add
  DatasetName <- "Mount Allison Phyto Lab Data"
  MonitoringLocationName <- "Kouchibouguac Bay"
  MonitoringLocationHorizontalCoordinateReferenceSystem <- "UNKWN"
  MonitoringLocationHorizontalAccuracyMeasure <- NA
  MonitoringLocationHorizontalAccuracyUnit <- NA
  MonitoringLocationType <- "Ocean"	
  ActivityType <- "Sample-Routine"	
  ActivityMediaName	<- "Ocean Water"
  ActivityEndDate	<- NA
  ActivityEndTime	<- NA
  ActivityDepthHeightMeasure <- "-0.2"	
  ActivityDepthHeightUnit	<- "m"
  SampleCollectionEquipmentName	<- "Water Bottle"
  CharacteristicName <- "Fecal Coliform"
  MethodSpeciation <- NA	
  ResultSampleFraction <- NA	
  ResultUnit <- "MPN"	
  ResultValueType	<- "Actual"
  ResultDetectionCondition <- NA	
  ResultDetectionQuantitationLimitMeasure <- NA	
  ResultDetectionQuantitationLimitUnit <- NA	
  ResultDetectionQuantitationLimitType <- NA	
  ResultStatusID <- NA	
  ResultComment <- NA	
  ResultAnalyticalMethodID <- NA	
  ResultAnalyticalMethodContext	<- NA
  ResultAnalyticalMethodName <- "Multiple Tube Test"	
  AnalysisStartDate	<- NA
  AnalysisStartTime	<- NA
  AnalysisStartTimeZone	<- NA
  LaboratoryName <- "Environment and Climate Change Canada Microbiology Lab"	
  LaboratorySampleID <- NA

#bind columns together
ECCCDataKouch <- cbind(ECCCDataKouch, DatasetName, MonitoringLocationName, MonitoringLocationHorizontalCoordinateReferenceSystem, MonitoringLocationHorizontalAccuracyMeasure, MonitoringLocationHorizontalAccuracyUnit, MonitoringLocationType, ActivityType, ActivityMediaName, ActivityEndDate, ActivityEndTime, ActivityDepthHeightMeasure, ActivityDepthHeightUnit, SampleCollectionEquipmentName, CharacteristicName, MethodSpeciation, ResultSampleFraction, ResultUnit, ResultValueType, ResultDetectionCondition, ResultDetectionQuantitationLimitMeasure, ResultDetectionQuantitationLimitUnit, ResultDetectionQuantitationLimitType, ResultStatusID, ResultComment, ResultAnalyticalMethodID, ResultAnalyticalMethodContext, ResultAnalyticalMethodName, AnalysisStartDate, AnalysisStartTime, AnalysisStartTimeZone, LaboratoryName, LaboratorySampleID)

```

#Reorder columns to agree with DataStream
```{r reorder columns}
col_order <- c("DatasetName", 
           "MonitoringLocationID",
           "MonitoringLocationName", 
           "MonitoringLocationLatitude",
           "MonitoringLocationLongitude",
           "MonitoringLocationHorizontalCoordinateReferenceSystem", 
           "MonitoringLocationHorizontalAccuracyMeasure", 
           "MonitoringLocationHorizontalAccuracyUnit", 
           "MonitoringLocationType", 
           "ActivityType", 
           "ActivityMediaName", 
           "ActivityStartDate",
           "ActivityStartTime",
           "ActivityEndDate", 
           "ActivityEndTime", 
           "ActivityDepthHeightMeasure", 
           "ActivityDepthHeightUnit", 
           "SampleCollectionEquipmentName", 
           "CharacteristicName", 
           "MethodSpeciation", 
           "ResultSampleFraction", 
           "ResultValue",
           "ResultUnit", 
           "ResultValueType", 
           "ResultDetectionCondition", 
           "ResultDetectionQuantitationLimitMeasure", 
           "ResultDetectionQuantitationLimitUnit", 
           "ResultDetectionQuantitationLimitType", 
           "ResultStatusID", 
           "ResultComment", 
           "ResultAnalyticalMethodID", 
           "ResultAnalyticalMethodContext", 
           "ResultAnalyticalMethodName", 
           "AnalysisStartDate", 
           "AnalysisStartTime", 
           "AnalysisStartTimeZone", 
           "LaboratoryName", 
           "LaboratorySampleID")

ECCCDataKouch <- ECCCDataKouch[, col_order]

```


## Add Shediac data 
## Subset data - filter for Shediac Bay > 2016-01-01 (Sector ID "NB-07-020-001"); 33 sites in total (some not active)
```{r subset data shediac}
ECCCData3 <- filter(ECCCData, Sector_Secteur == "NB-07-020-001" & Date >= "2016-01-01")

```

## Combine and re-format columns "Sector_Secteur" (i.e. NB-07-020-001) and "Site" (i.e. NB_0019) for both the data and site location files.
```{r}
ECCCData3$Site <- gsub('[NB_]', '', ECCCData3$Site)

ECCCData3$MonitoringLocationID <- paste(ECCCData3$Sector_Secteur, ECCCData3$Site)

#ECCCLocations$Site <- gsub('[NB_]', '', ECCCLocations$Site)

#ECCCLocations$MonitoringLocationID <- paste(ECCCLocations$Sector_Secteur, ECCCLocations$Site)

```

## Join two dataframes by matching "Sector_Secteur" and "Site" columns.
```{r join dataframes}
ECCCDataShediac <- ECCCData3 %>% 
    full_join(x=., y=ECCCLocations, by = c("MonitoringLocationID" = "MonitoringLocationID")) %>%
    filter(!is.na(Date))
```

## Select relevant columns - start with bacteria only, include temp and salinity data later on.
```{r column select}
ECCCDataShediac <- ECCCDataShediac %>%
  select(MonitoringLocationID, Latitude, Longitude, Date, FC_MPN_CF_NPP_100_mL) 

#select(MonitoringLocationID, Latitude, Longitude, Date, FC_MPN_CF_NPP_100_mL, Temp_C, Sal_PPT_PPM) 
```

## Rename current columns to DataStream format
```{r column rename}
ECCCDataShediac <- ECCCDataShediac %>%
  rename("ResultValue" = "FC_MPN_CF_NPP_100_mL", "MonitoringLocationLatitude" = "Latitude", "MonitoringLocationLongitude" = "Longitude", "ActivityStartDate" = "Date")

#Rename temp and salinity data later on:
#rename("Fecal Coliform" = "FC_MPN_CF_NPP_100_mL", "Temperature, water" = "Temp_C", "Salinity" = "Sal_PPT_PPM", "ActivityStartDate" = "Date")
```

## Use pivot longer only if including salinity and water temperature data; exclude at first to simplify
```{r pivot longer}
# ECCCDataAll %>%
#   pivot_longer(
#     #cols = c("Fecal Coliform", "Temperature, water", "Salinity"),
#     names_to = "CharacteristicName",
#     values_to = "ResultValue"
#   )

```

## Date format
Tell R to recognize the date column as "date"
```{r date format}
ECCCDataShediac <- tidyr::separate(ECCCDataShediac, ActivityStartDate, c("ActivityStartDate", "ActivityStartTime"), sep = " ") %>%
  mutate(ActivityStartDate = ymd(ActivityStartDate))
```

## Add all DataStream columns
Will need to repeat this chunk for each new location (i.e. "Shediac Bay") or parameter (i.e. temp and salinity) to be added. If adding many locations, write a function instead. For now, adding only 2 locations - this way okay.
```{r add columns}
#define new columns to add
  DatasetName <- "Mount Allison Phyto Lab Data"
  MonitoringLocationName <- "Shediac Bay"
  MonitoringLocationHorizontalCoordinateReferenceSystem <- "UNKWN"
  MonitoringLocationHorizontalAccuracyMeasure <- NA
  MonitoringLocationHorizontalAccuracyUnit <- NA
  MonitoringLocationType <- "Ocean"	
  ActivityType <- "Sample-Routine"	
  ActivityMediaName	<- "Ocean Water"
  ActivityEndDate	<- NA
  ActivityEndTime	<- NA
  ActivityDepthHeightMeasure <- "-0.2"	
  ActivityDepthHeightUnit	<- "m"
  SampleCollectionEquipmentName	<- "Water Bottle"
  CharacteristicName <- "Fecal Coliform"
  MethodSpeciation <- NA	
  ResultSampleFraction <- NA	
  ResultUnit <- "MPN"	
  ResultValueType	<- "Actual"
  ResultDetectionCondition <- NA	
  ResultDetectionQuantitationLimitMeasure <- NA	
  ResultDetectionQuantitationLimitUnit <- NA	
  ResultDetectionQuantitationLimitType <- NA	
  ResultStatusID <- NA	
  ResultComment <- NA	
  ResultAnalyticalMethodID <- NA	
  ResultAnalyticalMethodContext	<- NA
  ResultAnalyticalMethodName <- "Multiple Tube Test"	
  AnalysisStartDate	<- NA
  AnalysisStartTime	<- NA
  AnalysisStartTimeZone	<- NA
  LaboratoryName <- "Environment and Climate Change Canada Microbiology Lab"	
  LaboratorySampleID <- NA

#bind columns together
ECCCDataShediac <- cbind(ECCCDataShediac, DatasetName, MonitoringLocationName, MonitoringLocationHorizontalCoordinateReferenceSystem, MonitoringLocationHorizontalAccuracyMeasure, MonitoringLocationHorizontalAccuracyUnit, MonitoringLocationType, ActivityType, ActivityMediaName, ActivityEndDate, ActivityEndTime, ActivityDepthHeightMeasure, ActivityDepthHeightUnit, SampleCollectionEquipmentName, CharacteristicName, MethodSpeciation, ResultSampleFraction, ResultUnit, ResultValueType, ResultDetectionCondition, ResultDetectionQuantitationLimitMeasure, ResultDetectionQuantitationLimitUnit, ResultDetectionQuantitationLimitType, ResultStatusID, ResultComment, ResultAnalyticalMethodID, ResultAnalyticalMethodContext, ResultAnalyticalMethodName, AnalysisStartDate, AnalysisStartTime, AnalysisStartTimeZone, LaboratoryName, LaboratorySampleID)

```

## Put columns in DataStream order
```{r reorder columns}

col_order <- c("DatasetName", 
           "MonitoringLocationID",
           "MonitoringLocationName", 
           "MonitoringLocationLatitude",
           "MonitoringLocationLongitude",
           "MonitoringLocationHorizontalCoordinateReferenceSystem", 
           "MonitoringLocationHorizontalAccuracyMeasure", 
           "MonitoringLocationHorizontalAccuracyUnit", 
           "MonitoringLocationType", 
           "ActivityType", 
           "ActivityMediaName", 
           "ActivityStartDate",
           "ActivityStartTime",
           "ActivityEndDate", 
           "ActivityEndTime", 
           "ActivityDepthHeightMeasure", 
           "ActivityDepthHeightUnit", 
           "SampleCollectionEquipmentName", 
           "CharacteristicName", 
           "MethodSpeciation", 
           "ResultSampleFraction", 
           "ResultValue",
           "ResultUnit", 
           "ResultValueType", 
           "ResultDetectionCondition", 
           "ResultDetectionQuantitationLimitMeasure", 
           "ResultDetectionQuantitationLimitUnit", 
           "ResultDetectionQuantitationLimitType", 
           "ResultStatusID", 
           "ResultComment", 
           "ResultAnalyticalMethodID", 
           "ResultAnalyticalMethodContext", 
           "ResultAnalyticalMethodName", 
           "AnalysisStartDate", 
           "AnalysisStartTime", 
           "AnalysisStartTimeZone", 
           "LaboratoryName", 
           "LaboratorySampleID")

ECCCDataShediac <- ECCCDataShediac[, col_order]

```

## Combine reformatted Kouch and Shediac Bay 2016 - 2020 ECCC datasets 
In order to upload to DataStream would need to add line removing NA values - not required for internal use.
```{r bind}
ECCCDataKouchShediac <- rbind(ECCCDataKouch, ECCCDataShediac)

```

## Save as .Rds for later use

```{r save .Rds}
saveRDS(ECCCDataKouchShediac, file = file.path(TidiedData, paste("ECCCDataKouchShediac", ".Rds", sep = ""), fsep = .Platform$file.sep))
```

NOTE: the 2 MPN limit of detection for ECCC shellfish survey data was not recorded using the result detection quantitation limit columns in the "Mount Allison Phyto Lab Data" currently on DataStream. This breakdown was therefore not include here for consistency and also to avoid removing measurements from the analysis (i.e. data below reporting limit has a "ResultValue" of "blank" and the fact that a measurement was collected on a given day would be lost). While the limit columns are not required for the CSM 2023 poster data analysis, they should be added if this dataset is uploaded to DataStream. There is a code chunk for this formatting in Angela's/Doug's old .Rmd "DS_Conversion.Rmd".




