---
title: "WeatherCanImport"
author: 
- Miranda Corkum^1^
- Emma Sutherland^1^ 
- Naaman Omar^1^
- Douglas Campbell^1^*
date: "`r format(Sys.Date())`"
output: 
  bookdown::word_document2:
    toc: TRUE
  bookdown::html_document2:
    code_folding: show
    toc: TRUE
    toc_float: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: [RPackageCitations.bib, packages.bib]
---

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

*corresponding author

# Acknowledgements {-}
Brackets minus after heading excludes heading from numbering.  
MC was supported by the Canada Research Chairs.
ES was supported by the New Brunswick Wildlife Trust Fund and a Cleantech Internship.  
NO was supported by
DAC was supported by the Canada Research Chairs.


# Introduction

This script uses packages to support direct import of Environment and Climate Change Canada weather (and tide) data into R. New columns are added generating variables used in a Shiny app (Code/"Facet_app_WQ.R") for data exploration. Weather data are saved as TidiedData/"KouchWeatherDay.Rds", "ShediacWeatherDay.Rds" for further use in Code/"DataClean.Rmd".


# Materials and Methods

## Set chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Update Weathercan data *only needs to be done occassionally*
These packages need to be installed to update Weathercan data. They are installed separately because they are not part of the CRAN repository.
Update weathercan to 0.7.1 (September 20, 2023)
https://github.com/ropensci/weathercan
https://docs.ropensci.org/weathercan/
```{r install packages}
#install.packages("weathercan", 
                 #repos = c("https://ropensci.r-universe.dev", 
                           #"https://cloud.r-project.org"))
```


## Load libraries
Set up libraries (or 'packages') - these contain additions to base R
Install all the packages in the tidyverse by running install. packages("tidyverse"). Run library(tidyverse) to load the core tidyverse and make it available in your current R session.
Note - lubridate is now part of Tidyverse version 2 or later.
```{r load libraries, include = FALSE}
library(tidyverse)

library(weathercan)

#Package 'lutz' and its dependency, 'sf', are required to get timezones for the updated stations dataframe.
library(lutz)

```

## Citations of software packages
Analysis was done using the 'tidyverse' [@R-tidyverse], 'weathercan' [@R-weathercan] and 'lutz' [@R-lutz] packages. 
Keep this chunk commented out or it will overwrite the final manuscript version of the .bib file.
```{r cite packages}
#knitr::write_bib(c('tidyverse', 'weathercan', 'lutz'), 'packages.bib')
```

##Update Weathercan station data to get the most recent data
Requires 'lutz' package for timezones.
```{r Weathercan update}
#stations_dl()
```

## Project Variables  
Assign project-specific values to variables
*Processed data contains any data that has been modified to include new columns etc. It can also contain the output of many files merged together.
*Tidied data contains a clean version of the raw data (Correct file names, dates in the right format etc).
*Do not use hard file path - not cross-platform compatible - use R functions to construct...
```{r set variables}
Project <- "northumberlandwaterquality"
DataIn <- file.path("..", "DataIn")
ProcessedData <- "ProcessedData"
Figures <- file.path("..", "Figures")
TidiedData <- file.path("..", "TidiedData")

EnCanStation1 <- "KOUCHIBOUGUAC"
Latitude1 <- 46.79
Longitude1 <- -65.01

EnCanStation2 <-"MONCTON"
Latitude2 <- 46.11
Longitude2 <- 64.68

Start <- 20100101000000
End <-Sys.time()
#TrimStart <- 20200401000000
#TrimEnd <- 20200501000000
#TideTolerance <- 5
  
FileEncode <- "UTF-8"

Delimiter <- ","

HeaderRows <- 0

Comment <- '#'

```

## Search for Enviroment Canada weather station
```{r weather data station}
#search for StationNames containing the value of EnCanStation set above
StationNames1 <- weathercan::stations_search(EnCanStation1)
#StationNames1

stations_search("Kouchibouguac", interval = "day")
```

StationID `r as.character(StationNames$station_id[1])` based upon search for `r EnCanStation`.

```{r weather data station}
#search for StationNames containing the value of EnCanStation set above
StationNames2 <- weathercan::stations_search(EnCanStation2)
#StationNames2

stations_search("MONCTON", interval = "day")
```

## Download Environment Canada weather data - Kouchibouguac
ECCC has shellfish survey data for Kouchibouguac starting 1985 to 2020 (database last accessed May 2023).
P. Godin - has a data technician looking into 2021-23 data, as it should be posted almost to date.
*Currently have data from 1991 + 2000-2020 in DataStream dataset.
Access daily weather data for date range 2000 to present 
**24May2023 added data for 2015 to 2020 = add Station ID 54538 and bind.

Station ID 6187: 1972-1995
Station ID 54538: 2016-2023
Station ID 26968: 1995-2018

```{r weather import}
#Be careful; time is reported as AST without ADT 
#This weather station ID contains daily data ranging from 1995 to 2018: call data for 2001 to 2015 (change to 95 if dataset is expanded to include more data)
KouchWeatherDay2000 <- weathercan::weather_dl(station_ids = 26968, start =  "2000-01-01", end = "2015-12-31", interval = "day")

#Call data interval "hour": 

#KouchWeatherHour <- weathercan::weather_dl(station_ids = 26968, start =  "2000-01-01", end = "2015-12-31", interval = "hour")

#This weather station ID contains data ranging from 2016 to 2023
KouchWeatherDay2016 <- weathercan::weather_dl(station_ids = 54538, start =  "2016-01-01", end = "2022-12-31", interval = "day")

#This weather station ID contains data ranging from 1972 to 1995: call data from 1991 to 1994
#KouchWeatherDay2 <- weathercan::weather_dl(station_ids = 6187, start =  "1991-01-01", end = "1994-12-31", interval = "day")

#Combine 2 data frames with the same column names using rbind: 

KouchWeatherDay <- rbind(KouchWeatherDay2000, KouchWeatherDay2016)

```

#DC 4Apr2023 - A work around for: 
1) total precip - 24/48 hr lag and summing across days (48/72 hrs counting sample day)
2) spd_max_gust (wind) - 24/48 hr lag and average for 48/72 hrs
3) mean_temp - 24/48 hr lag and average "mean_temp" for 48/72 hrs

#No wind data available for Kouch in current date window

```{r set timezone add columns}
#force_tz is a lubridate function to set a timezone for daily measurements

KouchWeatherDay <- KouchWeatherDay %>%
  mutate(date = force_tz(ymd(date),"America/Halifax")) %>%
  mutate(precip_lag24 = lag(total_precip, n = 1L)) %>%
  mutate(precip_lag48 = lag(total_precip, n = 2L)) %>%
  mutate(precip_sum48 = total_precip + precip_lag24) %>%
  mutate(precip_prior48 = precip_lag24 + precip_lag48) %>%
  mutate(precip_sum72 = total_precip + precip_lag24 + precip_lag48) %>%
  mutate(spd_max_gust_lag24 = lag(spd_max_gust, n = 1L)) %>%
  mutate(spd_max_gust_lag48 = lag(spd_max_gust, n = 2L)) %>%
  mutate(spd_max_gust_AVG48 = (spd_max_gust + spd_max_gust_lag24)/2) %>%
  mutate(spd_max_gust_AVG72 = (spd_max_gust + spd_max_gust_lag24 + spd_max_gust_lag48)/3) %>%
  mutate(mean_temp_lag24 = lag(mean_temp, n = 1L)) %>%
  mutate(mean_temp_lag48 = lag(mean_temp, n = 2L)) %>%
  mutate(mean_temp_AVG48 = (mean_temp + mean_temp_lag24)/2) %>%
  mutate(mean_temp_prior48 = (mean_temp_lag24 + mean_temp_lag48)/2) %>%
  mutate(mean_temp_AVG72 = (mean_temp + mean_temp_lag24 + mean_temp_lag48)/3)
 
```

## Inspect the imported ECCC weather data  
head will show the first 6 rows
```{r inspect data}
head(KouchWeatherDay)

```

## Save Kouchibouguac weather data - quick access for further analysis
```{r save data}
saveRDS(KouchWeatherDay, file = file.path(TidiedData, paste("KouchWeatherDay", ".Rds", sep = ""), fsep = .Platform$file.sep))
                                                
```

## Download Environment Canada weather data - Shediac
ECCC has shellfish survey data for Shediac Bay (Parlee Beach Provincial Park) starting in 1988 to 2018 (database last accessed September 2021; updated every 3 years i.e. winter 2021). 
NB Dept of Health Recreational Beach Water Quality Monitoring website has data from 2017-2023. 
Access daily weather data for date range 1990-2023.
*Note - Parlee Beach and Shediac Bay share same weather station data.
```{r}
#This weather station ID contains data ranging from years 1939 to 2012 - call years 1990 to 2012: 
ShediacWeather1990Day <- weathercan::weather_dl(station_ids = 6207, start =  "1990-01-01", interval = "day")

#This weather station ID contains data ranging from years 2012 to 2023; call years 2012 to 2022: 
ShediacWeather2012Day <- weathercan::weather_dl(station_ids = 50309, start =  "2012-01-01", end = "2023-12-31", interval = "day")

#Combine 2 data frames with the same column names using rbind: 

# combine by row
ShediacWeatherDay <- rbind(ShediacWeather1990Day, ShediacWeather2012Day)

# output
#View(ShediacWeatherDay)
#write.csv(ShediacWeatherDay, file = file.path(DataIn, paste("ShediacWeatherDay", ".csv", sep = ""), fsep = .Platform$file.sep), row.names=FALSE)
   
```

```{r set timezone}
#force_tz is a lubridate function to set a timezone for daily measurements

ShediacWeatherDay <- ShediacWeatherDay %>%
  mutate(date = force_tz(ymd(date),"America/Halifax")) %>%
  mutate(precip_lag24 = lag(total_precip, n = 1L)) %>%
  mutate(precip_lag48 = lag(total_precip, n = 2L)) %>%
  mutate(precip_sum48 = total_precip + precip_lag24) %>%
  mutate(precip_prior48 = precip_lag24 + precip_lag48) %>%
  mutate(precip_sum72 = total_precip + precip_lag24 + precip_lag48) %>%
  mutate(spd_max_gust_lag24 = lag(spd_max_gust, n = 1L)) %>%
  mutate(spd_max_gust_lag48 = lag(spd_max_gust, n = 2L)) %>%
  mutate(spd_max_gust_AVG48 = (spd_max_gust + spd_max_gust_lag24)/2) %>%
  mutate(spd_max_gust_AVG72 = (spd_max_gust + spd_max_gust_lag24 + spd_max_gust_lag48)/3) %>%
  mutate(mean_temp_lag24 = lag(mean_temp, n = 1L)) %>%
  mutate(mean_temp_lag48 = lag(mean_temp, n = 2L)) %>%
  mutate(mean_temp_AVG48 = (mean_temp + mean_temp_lag24)/2) %>%
  mutate(mean_temp_prior48 = (mean_temp_lag24 + mean_temp_lag48)/2) %>%
  mutate(mean_temp_AVG72 = (mean_temp + mean_temp_lag24 + mean_temp_lag48)/3)
 
```

## Inspect the imported ECCC weather data  
head will show the first 6 rows
```{r inspect data}
head(ShediacWeatherDay)

```

## Save Shediac weather data - quick access for further analysis
```{r save data}
saveRDS(ShediacWeatherDay, file = file.path(TidiedData, paste("ShediacWeatherDay", ".Rds", sep = ""), fsep = .Platform$file.sep))
                                             
```

